---
title: "Connecting R and Tableau"
author: Kyle Monahan
date: 10-22-2018
output: html_notebook
---
#Background
This is an R Markdown notebook highlighting the R code you need to connect R and Tableau. 

This was completed using the following software: 

* RStudio 1.1.456 - Mozilla/5.0 (Macintosh; Intel Mac OS X 10_13_6) AppleWebKit/605.1.15 
* Tableau 10.5 

#Getting Started
To start working with R, we have to use `RServe`, which is an R server. This is a server that is listening for R commands to pass to the R kernel. The default port for this server is **6311**, but I found port **84** worked for me. 

The code to install (commented out with #) and run Rserver is below. Note that the argument parameter no-save is required on *nix (Linux, Mac) systems. If you're on Windows, you probably don't need this. 

```{r}

#install.packages("Rserve")
library(Rserve)
Rserve(args = "--no-save") 

```

#Opening RServe in Tableau 

Open Tableau and go to: `Help > Setting and Performance > Managing External Service Connections`. This is where you can connect to the RServer you just started. From there, you can load R locally by putting in localhost and port **6311**. Note that if you were running Rserve on a virtual machine, you could direct that here as well. 

Press Enter. If you see "Successfully connected to the RServe service" then you are all set. 

#Running an example batch of code

You can find more information at the links I provided below, but the key notes are: 

* You can use the following functions to bring things into and out of Tableau: 

* 


#Running this example in R
The code we are running in Tableau looks like: 

> `SCRIPT_STR("library(syuzhet);r<-rescale(get_sentiment(.arg1,method = 'syuzhet'));ifelse(abs(r) > 0.3, 'positive',  'negative')",ATTR([Tweet]))`

In R, this would look like: 

```{r}
#Set working directory
setwd("/Users/kmonah02/Desktop") #To my local directory
#Import data 
library(readr)
library(stringr)
arg1 <- read_csv("1377884570_tweet_global_warming.csv") #Doesn't have to be called arg1, I just named it that for consistency, please email kyle DOT monahan @ tufts.edu if you want to check out this sample file, or download from here: https://www.figure-eight.com/data-for-everyone/

#Clean the tweets
arg1$cleantweet=stringr::str_replace_all(arg1$tweet,"[^[:graph:]]", " ") 

#Load the data into the sentiment analysis machine (as a character vector)
library(syuzhet)
r<-rescale(get_sentiment(arg1$cleantweet,method = 'syuzhet'))
ifelse(test = abs(r) > 0.3, yes =  'positive',no = 'negative')

#Save out the data
write.csv(x = arg1,file = "1377884570_tweet_global_warming_clean.csv")


```



#More information
https://www.tableau.com/learn/whitepapers/using-r-and-tableau